---
title: "My Dendrogram"
author: "A. Paul"
date: "6 Januar 2018"
runtime: shiny
output:
  ioslides_presentation: 
    highlight: monochrome
    incremental: yes
    keep_md: yes
    smaller: yes
    transition: faster
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Presentation Dendrogram

In order to gain an initial overview of unknown data, it is often useful to cluster and visualize the clusters. A first possibility is hierarchical clustering and a presentation in a dendrogram.

- The following Shiny application is used to create a dendrogram to represent the hierarchical clustering of a data frame.
- Initially, a test set of 5 data sets is displayed, which illustrates the presentation.
- The application user has the option of uploading his data from a csv file.
- After uploading, the number of records will be displayed as "size".
- The uploaded data is clustered and displayed as a dendrogram on the right. The dendrogram can be rotated.
- The creation of the hierarchical cluster can be defined by changing the metric and the method of clustering.
- Optionally, the user can view the csv data below the dendrogram.

## Plot Dendrogram

```{r include=FALSE}
library( shiny )
library( rsconnect )
library( tidyverse )
library( ggdendro )
library( ggplot2 )
```

```{r eruptions}
  # Sidebar with a slider input for number of bins 
  sidebarLayout(
    sidebarPanel(
      
      fileInput("file1", "Choose CSV File for data",
                accept = c(
                  "text/csv",
                  "text/comma-separated-values,text/plain",
                  ".csv") ),
      checkboxInput("header", "with header", TRUE),  
      checkboxInput("display", "with data", FALSE), 
      textOutput("name"),
      textOutput("size"),
      
      ## tags$hr(),
      
      selectInput( "method", "Choose a cluster method:",
                   list( `Single-Link`   = c("single"),
                         `Complete-Link` = c("complete"),
                         `Average-Link`  = c("average"),
                         `Median-Link`   = c("median"),
                         `Centroid-Link` = c("centroid") ) ),
      
      selectInput( "metric", "Choose a distance metric:",
                   list( `Euclidian` = c("euclidean"),
                         `Manhattan` = c("manhattan"),                       
                         `Minkowski` = c("minkowski"),
                         `Binary`    = c("binary") ) ),
      
      numericInput( "dim", "Minkowski dimension", value=1 ),
      
      checkboxInput("rotate", "Rotate", TRUE)

    ),
    
    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("plotDendrogram"),
      tableOutput("data")
    )
  )

output$plotDendrogram <- renderPlot({  
    
  inFile <- input$file1
    
  if ( is.null(inFile) ) { 
    
##  name
##  output$name <- renderText( "test data" )
##  test data
    my_data <- data.frame( x = c( 2, 4, 5, 3, 1 ), 
                           y = c( 6, 9, 6, 2, 7 ) )
    rownames(my_data) <- c("one", "two", "three", "four", "five" )
 
  } else {
 
##  name
##  output$name <- renderText( paste0( "file : ", inFile$name ) )
##  read csv-file with matrix     
    my_data <- read.csv( inFile$datapath, header=input$header )
    
  } 

  ## output$name <- renderText( l_name )
    
  l_text <- paste0( "size: ", nrow(my_data) )  
  output$size <- renderText( l_text )
    
  if ( input$display == TRUE ) { 
    output$data <- renderTable( my_data, rownames = TRUE)
  } else {
    output$data <- renderText( paste0("") )
  }
  
  ## generate cluster dendogram based on input$metric and input$method
  mat.dist <- dist( x=my_data, method=input$metric, p=input$dim )
  ## generate hierarchical     
  h.clust  <- hclust( mat.dist, method=input$method )
  ggdendro::ggdendrogram( h.clust, rotate=input$rotate, size=4, theme_dendro = TRUE )
  
} )
```


## Explanation of the selection parameters

* Box for "Choose CSV File" :
  Via a dialog window, a csv file can be selected and uploaded.

* Checkbox for "CSV file with Header" :
  If the csv file has a header, the checkbox should be selected.

* Checkbox for "Display data" :
  The checkbox allows the user to display the data under the Dendrogram.

* Selection list for a "Choose a cluster method" :
  The following procedures are offered: "Single-Link", "Complete Link", "Average-Link", "Median-Link" or "Centroid link"

* Selection list for a "Choose a distance metric" :
  The following metrics are offered: "Euclidean", "Manhattan", "Minkowski" or "Binary"

* For the Minkowski metric an additional dimension can be chosen.


## Test Data and Clustering

To illustrate, here are the test data.
```{r echo=FALSE}
  ##  test data
  my_data <- data.frame( x = c( 2, 4, 5, 3, 1 ), 
                         y = c( 6, 9, 6, 2, 7 ) )
  rownames(my_data) <- c("one", "two", "three", "four", "five" )
  ## my_data
  print( my_data )
```

```{r}
  ## generate cluster dendogram based on input$metric and input$method
  l_metric <- "manhattan"
  mat.dist <- stats::dist( x=my_data, method=l_metric )
  ## generate hierarchical
  l_method <- "single"
  h.clust <- stats::hclust( mat.dist, method=l_method )
  ## dendrogram
  my_dend <- stats::as.dendrogram(h.clust)
  ## print(h.clust)
  paste0( "Cluster method : ", h.clust$method, " with ", "Distance : ", h.clust$dist.method )
```

Enclosed the clustering with the leaves and the gradual union including the heights.
```{r}
str(my_dend)
```


